Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-10-13 grove-paper.tex
start math for edit action syntax and semantics (Eric)

% math ligatures
\mathlig{::=}{\coloncolonequals}
\mathlig{|-}{\vdash}
\mathlig{:->}{\mapsto}
\mathlig{-->a}{\overset{a}{\longrightarrow}}
%\mathlig{||}{|}
%\mathlig{|}{:}
%\mathlig{:}{\!:\!}
\newcommand{\abs}[1]{\left|{#1}\right|}

@@

@@

% sets
\def\A{\mathcal{A}}
\def\C{\mathcal{C}}
\def\E{\mathcal{E}}
\def\G{\mathcal{G}}
\def\I{\mathcal{I}}
\def\K{\mathcal{K}}
\def\P{\mathcal{P}}
\def\S{\mathcal{S}}
\def\U{\mathcal{U}}
\def\V{\mathcal{V}}
\def\e{\varepsilon}
\def\AA{\textbf{A}}
\def\EE{\textbf{E}}

@@

% relations
\DeclareMathOperator{\arity}{\text{arity}}
\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
\DeclareMathOperator{\deleted}{\text{deleted}}
\DeclareMathOperator{\dom}{\text{dom}}
\DeclareMathOperator{\edges}{\text{edges}}
\DeclareMathOperator{\known}{\text{known}}
\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
\DeclareMathOperator{\multiparents}{\text{multiparents}}
\DeclareMathOperator{\orphans}{\text{orphans}}
\DeclareMathOperator{\parents}{\text{parents}}
\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
\DeclareMathOperator{\reachable}{\text{reachable}}
\DeclareMathOperator{\seen}{\text{seen}}
\DeclareMathOperator{\unseen}{\text{unseen}}
\DeclareMathOperator{\vertexes}{\text{vertexes}}
\DeclareMathOperator{\vertex}{\text{vertex}}
\DeclareMathOperator{\publish}{\text{publish}}
\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
\DeclareMathOperator{\downIndex}{\text{downIndex}}
\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}

@@

You can map from syntax to graph by selecting
unique IDs (for both edges and vertexes).
    - holes are not explicit in the graph

@@

@@

\begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
  $\A$ & Actions & $\S$ & Edge States   \\
  $\C$ & Cursors & $\U$ & UUIDs         \\
  $\E$ & Edges   & $\V$ & Vertexes      \\
  $\G$ & Graphs  & $\AA$ & Graph Actions \\
  $\I$ & Indices & $\EE$ & Editors       \\
  $\K$ & Constructors                   \\
\end{tabular}

@@

A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$. Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference to all of the rooted vertexes, of a graph.

@@

An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.

@@

A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.

@@

An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \EE = \mathcal{G} \times \C \times \AA^{*} \times \AA^{*} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.

>>>>

% math ligatures
\mathlig{::=}{\coloncolonequals}
\mathlig{|-}{\vdash}
\mathlig{:->}{\mapsto}
\mathlig{-->a}{\overset{a}{\longrightarrow}}
\newcommand{\abs}[1]{\left|{#1}\right|}

@@

\newcommand{\conflict}[2]{\llparenthesis#1|#2\rrparenthesis}

@@

% sets
\def\A{\mathcal{A}}
%\def\C{\mathcal{C}}
\def\E{\mathcal{E}}
\def\G{\mathcal{G}}
\def\I{\mathcal{I}}
\def\K{\mathcal{K}}
\def\P{\mathcal{P}}
\def\U{\mathcal{U}}
\def\V{\mathcal{V}}
\def\e{\varepsilon}
\def\AA{\textbf{A}}
\def\EE{\textbf{E}}

@@

% relations
\DeclareMathOperator{\s}{\text{S}}
\DeclareMathOperator{\MP}{\text{MP}}
\DeclareMathOperator{\CC}{\text{C}}
\DeclareMathOperator{\D}{\text{D}}

\DeclareMathOperator{\arity}{\text{arity}}
\DeclareMathOperator{\childVertexes}{\text{childVertexes}}
\DeclareMathOperator{\cursorChildren}{\text{cursorChildren}}
\DeclareMathOperator{\vertexChildren}{\text{vertexChildren}}
\DeclareMathOperator{\deleted}{\text{deleted}}
\DeclareMathOperator{\dom}{\text{dom}}
\DeclareMathOperator{\edges}{\text{edges}}
\DeclareMathOperator{\known}{\text{known}}
\DeclareMathOperator{\liveEdges}{\text{liveEdges}}
\DeclareMathOperator{\liveVertexes}{\text{liveVertexes}}
\DeclareMathOperator{\multiparents}{\text{multiparents}}
\DeclareMathOperator{\orphans}{\text{orphans}}
\DeclareMathOperator{\parents}{\text{parents}}
\DeclareMathOperator{\parentVertexes}{\text{parentVertexes}}
\DeclareMathOperator{\reachable}{\text{reachable}}
\DeclareMathOperator{\seen}{\text{seen}}
\DeclareMathOperator{\unseen}{\text{unseen}}
\DeclareMathOperator{\vertexes}{\text{vertexes}}
\DeclareMathOperator{\vertex}{\text{vertex}}
\DeclareMathOperator{\publish}{\text{publish}}
\DeclareMathOperator{\leftIndex}{\text{leftIndex}}
\DeclareMathOperator{\rightIndex}{\text{rightIndex}}
\DeclareMathOperator{\downIndex}{\text{downIndex}}
\DeclareMathOperator{\defaultIndex}{\text{defaultIndex}}

@@

You can map from syntax to graph by selecting
unique IDs $u$ (for both edges and vertexes).
    - holes are not explicit in the graph

@@

\[
\begin{array}{l}
  \textrm{Edit Actions}: \grave{e} \in Act ::= e \mid \#u \mid \conflict{\grave{e}}{\grave{e}} \\
  \s : \G \to Act \times \V^{*} \times \V^{*} \times \V^{*} \\
  \s(G) = (\grave{e}, \MP(G), \CC(G), \D(G)) \\
  \MP(G) = \left\{ v \in \vertexes(G) \mid \exists e_1,e_2 \in \liveEdges(G), e_1= \right. \\
      \qquad \left. (v,p_1,v_1), e_2=(v,p_2,v_2), e_1 \ne e_2 \right\} \\
  \CC(G) = \ldots \\
  \D(G) = \orphans(G) \setminus \left\{ v_0 \right\} \\
  \orphans(G) = \left\{ v \in \vertexes(G) \mid \forall e=(v^\prime,p,v^{\prime\prime}) \in \right. \\
      \qquad \left. \liveEdges(G), v^{\prime\prime} \ne v \right\} \\
  \vertexes(G) = \left\{ v \mid \exists e = (v^\prime, p, v^{\prime\prime}) \in \edges(G), v \in \left\{ v^\prime, v^{\prime\prime}\right\} \right\} \\
  \edges(G) = \left\{ e \mid G(e) \ne \bot \right\} \\
  \liveEdges(G) = \left\{ e \mid G(e) = + \right\} \\
\end{array}
\]

@@

\begin{tabular}{cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl@{\hspace{1.5cm}}cl}
  $\A$ & Actions & $\S$ & Edge States   \\
%  $\C$ & Cursors & $\U$ & UUIDs         \\
  $\E$ & Edges   & $\V$ & Vertexes      \\
  $\G$ & Graphs  & $\AA$ & Graph Actions \\
  $\I$ & Indices & $\EE$ & Editors       \\
  $\K$ & Constructors                   \\
\end{tabular}

@@

%A \emph{cursor} $c = (v,i) \in \C = \V \times \I$ is a reference to the (possibly empty) set of all vertexes with parent vertex $v$ and child index $i$. Let $c_{root} = (v_{root}, i_{root})$ be the \emph{root cursor}, a reference to all of the rooted vertexes, of a graph.

@@

%An \emph{edge} $\e = (c, v, u) \in \E = \C \times \V \times \U$ adds vertex $v$ with the set of vertexes referenced by cursor $c$.

@@

%A \emph{move action} $a \in \A^{move} = \{ \Left,\Right,\Up,\Down \} \cup \{ \Select \} \times \C$ re-positions the cursor.

@@

%An \emph{editor} $E = (G,c,A_{Q}^{*},A_{H}^{*},u) \in \EE = \mathcal{G} \times \C \times \AA^{*} \times \AA^{*} \times \U$ is an instance of graph $G$ with cursor $c$, queued graph actions $A_{Q}^{*}$, and known graph action history $A_{H}^{*}$, that can be identified by UUID $u$.
